---
title: "2022b MiSeq SCRuB of Plate 1"
author: "Brooke Benz"
date: "2024-06-10"
output:
  html_notebook: 20240610_BarleyMicrobiome_SCRuB_2022b_miseq_BRB
  pdf_document: 20240610_BarleyMicrobiome_SCRuB_2022b_miseq_BRB
---


* Run using `r version[['version.string']] `.
---

#### Load Packages, set path to data (hidden)
```{r, echo=FALSE, results='hide', include=FALSE} 
#install.packages("glmnet") #when installing glmnet input no for do you wnat to install from sources the package which needs compilation
#install.packages("torch") #when installing select Yes when prompted

x<-c("ggplot2", "phyloseq", "Biostrings", "dplyr", "tidyverse", "glmnet", "torch")  
lapply(x, require, character.only = TRUE)

#require(devtools)
#devtools::install_github("shenhav-and-korem-labs/SCRuB")
#require(SCRuB)
library(SCRuB)

#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
set.seed(617)

# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

path <- "/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB"

# load merged phyloseq object 
ps_22b_ms <- readRDS("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/Merging PS Objects/2022b_Plate1_miseq.ps.rds")
ps_22b_ms
```
## Changing the metadata files
```{r}
metadata_22b_ms <- (as(sample_data(ps_22b_ms), "data.frame"))
metadata_22b_ms
```

```{r}
#Update column ControlType to Metadata file
metadata_22b_ms <- metadata_22b_ms %>%
  mutate(ControlType = ifelse(SampleID == "P1-Ex-Neg-Ctrl", "ExNormalization", as.character(ControlType)))
metadata_22b_ms <- metadata_22b_ms %>%
  mutate(ControlType = ifelse(SampleID == "P1-Ex-Neg-Ctrln", "ExNormalization", as.character(ControlType)))

metadata_22b_ms <- metadata_22b_ms %>%
  mutate(ControlType = ifelse(SampleID == "P1-PCR-Neg-Ctrl", "PCRNormalization", as.character(ControlType)))
metadata_22b_ms <- metadata_22b_ms %>%
  mutate(ControlType = ifelse(SampleID == "P1-PCR-Neg-Ctrln", "PCRNormalization", as.character(ControlType)))

metadata_22b_ms
```

```{r}
sample_data(ps_22b_ms) <- metadata_22b_ms

write.csv(metadata_22b_ms, file = "2022b_miseq_SCRuBmetadata.csv", row.names = FALSE)
```

## Phase1 SCRuB - PCR Neg Ctrl
### 2022b
```{r}
meta_phase1 <- (as(sample_data(ps_22b_ms), "data.frame"))
mat_phase1 <- (as(otu_table(ps_22b_ms), "matrix"))

# format for SCRuB
meta_phase1 <- meta_phase1 %>% arrange(SampleID) %>% #puts everything in order by sampleID
    select(is_control, sample_type, sample_well, sample_plate, original_plate, LibPrepMethod)#only selects these columns
meta_phase1$sample_well <- as.character(meta_phase1$sample_well)#converts sample_well to character variable

# rearrange rows to match - first order it by sample ID then remove sample id column
mat_phase1 %>% data.frame %>% mutate(ids = rownames(mat_phase1)) %>%
    arrange(ids) %>% select(-ids) %>% as.matrix -> mat_phase1

# subset by plate
meta_p1_o <- meta_phase1 %>% filter(LibPrepMethod == "Old") %>% select(is_control, sample_type, sample_well)
mat_p1_o <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p1_o),]
meta_p1_n <- meta_phase1 %>% filter(LibPrepMethod == "New") %>% select(is_control, sample_type, sample_well)
mat_p1_n <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p1_n),]

#Check that samples match and number of samples is correct - all should be TRUE and equal 96
rownames(mat_p1_o) %in% rownames(meta_p1_o) #96
rownames(mat_p1_n) %in% rownames(meta_p1_n) #96
```

### Run SCRuB
* Running SCRuB with spatial (well location) information. Done by plate.
```{r}
#scrub DNA plates (individually) by DNA extraction controls
scr_p1_o <- SCRuB(mat_p1_o, meta_p1_o, c("PCR-Neg-Ctrl") )
scr_p1_n <- SCRuB(mat_p1_n, meta_p1_n, c("PCR-Neg-Ctrl") ) 
```
## Saving and loading Scrubbed files
```{r}
#save(scr_p1_o, file="scrubOut_p1_oldlibprep.RData")
#save(scr_p1_n, file="scrubOut_p1_newlibprep.RData")

load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p1_oldlibprep.RData")
load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p1_newlibprep.RData")
```

## Examine output
### 2022b Plate 1 Old Library Prep Method
```{r}
# Estimated contamination level
scr_p1_o$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1_o$p*100))  #1.24% contamination
```
### 2022b Plate 1 New Library Prep method
```{r}
scr_p1_n$p %>% boxplot()
```
```{r}
median(100-(scr_p1_n$p*100))  #0.096% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1 Old Library Prep Method
```{r}
#  index = 1 to 4669 ASVs in this dataset
scr_p1_o$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_n$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```

## Estimate well-to-well leakage
### 2022b Plate 1 Old Library Prep Method
```{r}
scr_p1_o$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1_o$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.98% well-to-well leakage in plate 1 old library prep method
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_n$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1_n$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0% well leakage in plate 1 new library prep method
```

```{r}
decon_p1_o <- scr_p1_o$decontaminated_samples
decon_p1_n <- scr_p1_n$decontaminated_samples
write.csv(decon_p1_o, "2022b_miseq_ScrubDecontaminatedSamples_p1_oldlibprep_pcr.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p1_n, "2022b_miseq_ScrubDecontaminatedSamples_p1_newlibprep_pcr.csv")

#decon_p1_o <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_ScrubDecontaminatedSamples_p1_oldlibprep_pcr.csv", row.names = 1) %>% as.matrix()
#decon_p2_n <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_ScrubDecontaminatedSamples_p1_newlibprep_pcr.csv", row.names = 1) %>% as.matrix()
```

## Phase2 SCRuB - Extraction Neg Ctrl
### Run SCRuB
* Running SCRuB with spatial (well location) information. Done by plate.
```{r}
#scrub DNA plates (individually) by DNA extraction controls
scr_p1_ex_o <- SCRuB(mat_p1_o, meta_p1_o, c("Ex-Neg-Ctrl") )
scr_p1_ex_n <- SCRuB(mat_p1_n, meta_p1_n, c("Ex-Neg-Ctrl") ) 
```
## Saving and loading Scrubbed files
```{r}
#save(scr_p1_ex_o, file="scrubOut_p1_ex_oldlibprep.RData")
#save(scr_p1_ex_n, file="scrubOut_p1_ex_newlibprep.RData")

load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p1_ex_oldlibprep.RData")
load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p1_ex_newlibprep.RData")
```

## Examine output
### 2022b Plate 1 Old Library Prep Method
```{r}
# Estimated contamination level
scr_p1_ex_o$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1_ex_o$p*100))  #2.09% 
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_ex_n$p %>% boxplot()
```
```{r}
median(100-(scr_p1_ex_n$p*100))  #0.49% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1 Old Library Prep Method
```{r}
#  index = 1 to 7781 ASVs in this dataset
scr_p1_ex_o$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_ex_n$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```

## Estimate well-to-well leakage
### 2022b Plate 1 Old Library Prep Method
```{r}
scr_p1_ex_o$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_ex_o$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.94% well-to-well leakage in plate 1
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_ex_n$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_ex_n$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.78% well leakage in plate 2
```

```{r}
decon_p1_ex_o <- scr_p1_ex_o$decontaminated_samples
decon_p1_ex_n <- scr_p1_ex_n$decontaminated_samples
write.csv(decon_p1_ex_o, "2022b_miseq_ExNegCtrl_ScrubDecontaminatedSamples_p1_oldlibprep.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p1_ex_n, "2022b_miseq_ExNegCtrl_ScrubDecontaminatedSamples_p1_newlibprep.csv")

#decon_p1_ex_o <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_ExNegCtrl_ScrubDecontaminatedSamples_p1_oldlibprep.csv", row.names = 1) %>% as.matrix()
#decon_p1_ex_n <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_ExNegCtrl_ScrubDecontaminatedSamples_p1_newlibprep.csv", row.names = 1) %>% as.matrix()
```

# Output post-SCRuB 
```{r}
# create main datsets - putting plates back together again - combining PCR and Ex SCRuB
mat_final <- rbind(decon_p1_o, decon_p1_n)
ps_curated_0 <- subset_samples(ps_22b_ms, sample_names(ps_22b_ms) %in% rownames(mat_final))
ps_curated <- ps_curated_0
mat_table_final <- otu_table(mat_final, taxa_are_rows = FALSE)
otu_table(ps_curated) <- mat_table_final

mat_final_ex <- rbind(decon_p1_ex_o, decon_p1_ex_n)
ps_curated_0_ex <- subset_samples(ps_22b_ms, sample_names(ps_22b_ms) %in% rownames(mat_final_ex))
ps_curated_ex <- ps_curated_0_ex
mat_table_final_ex <- otu_table(mat_final_ex, taxa_are_rows = FALSE)
otu_table(ps_curated_ex) <- mat_table_final_ex

ps_curated_0@otu_table["BM-249",1:10] #prior SCRuB
```

## Seeing how ASV reads change per sample before and after scrubbing
### Plate 1 Old
```{r}
ps_curated@otu_table["BM-249",1:10]   #post SCRuB
```

### Plate 1 Old
```{r}
ps_curated_0@otu_table["BM-34",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-34",1:10]   #post SCRuB
```

### Plate 1 New
```{r}
ps_curated_0@otu_table["BM-249n",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-249n",1:10]   #post SCRuB
```

### Plate 1 New
```{r}
ps_curated_0@otu_table["BM-34n",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-34n",1:10]   #post SCRuB
```

### Plate 1 Old Ex
```{r}
ps_curated_0_ex@otu_table["BM-249",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-249",1:10]   #post SCRuB
```

### Plate 1 Old Ex
```{r}
ps_curated_0_ex@otu_table["BM-34",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-34",1:10]   #post SCRuB
```

### Plate 1 New Ex
```{r}
ps_curated_0_ex@otu_table["BM-249n",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-249n",1:10]   #post SCRuB
```

### Plate 1 New Ex
```{r}
ps_curated_0_ex@otu_table["BM-34n",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-34n",1:10]   #post SCRuB
```

```{r}
row_sums <- rowSums(ps_curated_0@otu_table) #prior scrub
print(row_sums)
```
```{r}
row_sums1 <- rowSums(ps_curated@otu_table) #post scrub
print(row_sums1)
```

```{r}
row_sums2 <- rowSums(ps_curated_0_ex@otu_table) #prior scrub
print(row_sums2)
```
```{r}
row_sums3 <- rowSums(ps_curated_ex@otu_table) #post scrub
print(row_sums3)
```

```{r}
# filter out taxa reduced to 0 reads (1305 ASVs, leaves 3364 out of 4669)
ps_curated <- filter_taxa(ps_curated, function(x) sum(x) > 0, TRUE)
#ps_22b_ms
ps_curated

# filter out taxa reduced to 0 reads (1325 ASVs, leaves 3344 out of 4669)
ps_curated_ex <- filter_taxa(ps_curated_ex, function(x) sum(x) > 0, TRUE)
ps_curated_ex
```

### Remove positive control samples
```{r}
ps_curated <- prune_samples(sample_names(ps_curated) != "P1-Pos-Ctrl", ps_curated)
ps_curated <- prune_samples(sample_names(ps_curated) != "P1-Pos-Ctrln", ps_curated)
ps_curated@sam_data
```
```{r}
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P1-Pos-Ctrl", ps_curated_ex)
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P1-Pos-Ctrln", ps_curated_ex)
ps_curated_ex@sam_data
```

## Merging SCRuBBED 2022 plates
### Merging Ex Neg Ctrl and PCR Neg Ctrl SCRuB
```{r}
#this will merge the samples with the same name - they will be added together
ps_22b_merge <- merge_phyloseq(ps_curated, ps_curated_ex)
ps_22b_merge
sample_data(ps_22b_merge)
#this combines PCR and Ex neg control scrub results, basically negating SCRuB
#will re-run together as a combined scrub
```

## 2022b Combined SCRuB
```{r}
## Running SCRuB of both Ex Neg and PCR Neg Ctrl at one time
## According to SCRuB tutorial, run SCRuB of first instance of contamination first

#scrub DNA plates (individually) by DNA extraction and PCR controls
scr_p1_comb_o <- SCRuB(mat_p1_o, meta_p1_o, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") )
scr_p1_comb_n <- SCRuB(mat_p1_n, meta_p1_n, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") )
```

## Saving and loading Scrubbed files
```{r}
#save(scr_p1_comb_o, file="scrubOut_p1_comb_oldlibprep.RData")
#save(scr_p1_comb_n, file="scrubOut_p2_comb_newlibprep.RData")

load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p1_comb_oldlibprep.RData")
load("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/scrubOut_p2_comb_newlibprep.RData")
```

## Examine output
### 2022b Plate 1 Old Library Prep Method
```{r}
# Estimated contamination level
scr_p1_comb_o$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1_comb_o$p*100))  #2.21% 
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_comb_n$p %>% boxplot()
```
```{r}
median(100-(scr_p1_comb_n$p*100))  #0.57% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1 Old Library Prep Method
```{r}
#  index = 1 to 7781 ASVs in this dataset
scr_p1_comb_o$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot()
scr_p1_comb_o$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_comb_n$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot()
scr_p1_comb_n$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```

## Estimate well-to-well leakage
### 2022b Plate 1 Old Library Prep Method
```{r}
scr_p1_comb_o$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_comb_o$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.94% well-to-well leakage in plate 1
scr_p1_comb_o$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1_comb_o$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.98% well-to-well leakage in plate 1
```
### 2022b Plate 1 New Library Prep Method
```{r}
scr_p1_comb_n$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_comb_n$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.78% well leakage in plate 2
scr_p1_comb_n$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1_comb_n$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0% well leakage in plate 2
```

```{r}
decon_p1_comb_o <- scr_p1_comb_o$decontaminated_samples
decon_p1_comb_n <- scr_p1_comb_n$decontaminated_samples
write.csv(decon_p1_comb_o, "2022b_miseq_comb_NegCtrl_ScrubDecontaminatedSamples_p1_oldlibprep.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p1_comb_n, "2022b_miseq_comb_NegCtrl_ScrubDecontaminatedSamples_p1_newlibprep.csv")

#decon_p1_comb_o <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_comb_NegCtrl_ScrubDecontaminatedSamples_p1_oldlibprep.csv", row.names = 1) %>% as.matrix()
#decon_p1_comb_n <- read.csv("/Users/brookeb/Documents/R/Final Codes/2022/2022_Miseq/SCRuB/2022b_miseq_comb_NegCtrl_ScrubDecontaminatedSamples_p1_newlibprep.csv", row.names = 1) %>% as.matrix()
```

# Output post-SCRuB 
```{r}
# create main datsets - putting plates back together again - combining PCR and Ex SCRuB
mat_final_comb <- rbind(decon_p1_comb_o, decon_p1_comb_n)
ps_curated_0_comb <- subset_samples(ps_22b_ms, sample_names(ps_22b_ms) %in% rownames(mat_final_comb))
ps_curated_comb <- ps_curated_0_comb
mat_table_final_comb <- otu_table(mat_final_comb, taxa_are_rows = FALSE)
otu_table(ps_curated_comb) <- mat_table_final_comb
```

### Plate 1 Combined Old
```{r}
ps_curated_0_comb@otu_table["BM-249",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-249",1:10]   #post SCRuB
```

### Plate 1 Combined Old
```{r}
ps_curated_0_comb@otu_table["BM-34",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-34",1:10]   #post SCRuB
```

### Plate 1 Combined New
```{r}
ps_curated_0_comb@otu_table["BM-249n",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-249n",1:10]   #post SCRuB
```

### Plate 1 Combined New
```{r}
ps_curated_0_comb@otu_table["BM-34n",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-34n",1:10]   #post SCRuB
```

```{r}
row_sums4 <- rowSums(ps_curated_0_comb@otu_table) #prior scrub
print(row_sums4)
```
```{r}
row_sums5 <- rowSums(ps_curated_comb@otu_table) #post scrub
print(row_sums5)
```

```{r}
# filter out taxa reduced to 0 reads (1938 ASVs, leaves 2731 out of 4669)
ps_curated_comb <- filter_taxa(ps_curated_comb, function(x) sum(x) > 0, TRUE)
#ps_22b_ms
ps_curated_comb
```

### Removing Positive Controls
```{r}
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P1-Pos-Ctrl", ps_curated_comb)
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P1-Pos-Ctrln", ps_curated_comb)
ps_curated_comb@sam_data
```

## Saving Files
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps_curated_comb))
names(dna) <- taxa_names(ps_curated_comb)
ps_22b_scrub <- merge_phyloseq(ps_curated_comb, dna)
taxa_names(ps_22b_scrub) <- paste0("ASV", seq(ntaxa(ps_22b_scrub)))
ps_22b_scrub

saveRDS(ps_22b_scrub,file.path(path,"2022b_miseq.ps.rds"))
```

```{r}
# saving files
scrub_meta <- (as(sample_data(ps_22b_scrub), "data.frame"))
scrub_taxa <- (as(otu_table(ps_22b_scrub), "matrix"))
# # write out data
write.csv(scrub_meta, "2022b_miseq_metadata_postSCRuB_comb.csv")
write.csv(scrub_taxa, "2022b_miseq_taxatable_postSCRuB_nextseq_comb.csv")
```

```{r}
sessionInfo()
```





###### end