---
title: "2022b NextSeq SCRuB"
author: "Brooke Benz"
date: "2024-05-15"
output:
  html_notebook: 20240515_BarleyMicrobiome_SCRuB_2022b_nextseq_BRB
  pdf_document: 20240515_BarleyMicrobiome_SCRuB_2022b_nextseq_BRB
---
## Contributors for this code included Brooke R. Benz NDSU and Briana K. Whitaker USDA-ARS

## Purpose: To remove contaminating sequences from samples using the SCRuB package.

* Run using `r version[['version.string']] `.
---
#### Load Packages, set path to data (hidden)
```{r, echo=FALSE, results='hide', include=FALSE} 
x<-c("ggplot2", "phyloseq", "Biostrings", "dplyr", "tidyverse", "glmnet", "torch")  
lapply(x, require, character.only = TRUE)

library(SCRuB)

#add 'not in' function
`%nin%` = Negate(`%in%`)

#set seed
set.seed(617)

# set ggplot2 theme
theme_set(theme_bw(base_size=16)) 
theme_update(panel.grid.major=element_line(0), panel.grid.minor=element_line(0))

path <- "~/Documents/R/Final Codes/2022/2022_Nextseq/SCRuB"

# load merged phyloseq object 
ps_22b <- readRDS("~/Documents/R/Final Codes/2022/2022_Nextseq/Merging PS Objects/2022b_trim251.ps.rds")
ps_22b
```
## Changing the metadata files
```{r}
metadata_22b <- (as(sample_data(ps_22b), "data.frame"))
metadata_22b
```

```{r}
#Update column ControlType to Metadata file
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P1-Ex-Neg-Ctrl", "ExNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P2-Ex-Neg-Ctrl", "ExNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P3-Ex-Neg-Ctrl", "ExNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P4-Ex-Neg-Ctrl", "ExNormalization", as.character(ControlType)))

metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P1-PCR-Neg-Ctrl", "PCRNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P2-PCR-Neg-Ctrl", "PCRNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P3-PCR-Neg-Ctrl", "PCRNormalization", as.character(ControlType)))
metadata_22b <- metadata_22b %>%
  mutate(ControlType = ifelse(SampleID == "P4-PCR-Neg-Ctrl", "PCRNormalization", as.character(ControlType)))

metadata_22b <- metadata_22b %>%
  mutate(sample_type = ifelse(is.na(sample_type), "seed", as.character(sample_type)))

metadata_22b
```

```{r}
sample_data(ps_22b) <- metadata_22b

write.csv(metadata_22b, file = "2022b_SCRuBmetadata.csv", row.names = FALSE)
```

## Phase1 SCRuB - PCR Neg Ctrl
### 2022b
```{r}
meta_phase1 <- (as(sample_data(ps_22b), "data.frame"))
mat_phase1 <- (as(otu_table(ps_22b), "matrix"))

# format for SCRuB
meta_phase1 <- meta_phase1 %>% arrange(SampleID) %>% #puts everything in order by sampleID
    select(is_control, sample_type, sample_well, sample_plate, original_plate)#only selects these columns
meta_phase1$sample_well <- as.character(meta_phase1$sample_well)#converts sample_well to character variable

# rearrange rows to match - first order it by sample ID then remove sample id column
mat_phase1 %>% data.frame %>% mutate(ids = rownames(mat_phase1)) %>%
    arrange(ids) %>% select(-ids) %>% as.matrix -> mat_phase1

# subset by plate
meta_p1 <- meta_phase1 %>% filter(sample_plate == "1") %>% select(is_control, sample_type, sample_well)
mat_p1 <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p1),]
meta_p2 <- meta_phase1 %>% filter(sample_plate == "2") %>% select(is_control, sample_type, sample_well)
mat_p2 <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p2),]
meta_p3 <- meta_phase1 %>% filter(sample_plate == "3") %>% select(is_control, sample_type, sample_well)
mat_p3 <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p3),]
meta_p4 <- meta_phase1 %>% filter(sample_plate == "4") %>% select(is_control, sample_type, sample_well)
mat_p4 <- mat_phase1[rownames(mat_phase1) %in% rownames(meta_p4),]

#Check that samples match and number of samples is correct - all should be TRUE and equal 96
rownames(mat_p1) %in% rownames(meta_p1) #96
rownames(mat_p2) %in% rownames(meta_p2) #96
rownames(mat_p3) %in% rownames(meta_p3) #96
rownames(mat_p4) %in% rownames(meta_p4) #68 - not a full plate
```

### Run SCRuB
* Running SCRuB with spatial (well location) information. Done by plate.
```{r}
#scrub DNA plates (individually) by DNA extraction controls
scr_p1 <- SCRuB(mat_p1, meta_p1, c("PCR-Neg-Ctrl") )
scr_p2 <- SCRuB(mat_p2, meta_p2, c("PCR-Neg-Ctrl") ) 
scr_p3 <- SCRuB(mat_p3, meta_p3, c("PCR-Neg-Ctrl") )
scr_p4 <- SCRuB(mat_p4, meta_p4, c("PCR-Neg-Ctrl") )
```
## Saving and loading Scrubbed files
```{r}
#save(scr_p1, file="scrubOut_p1.RData")
#save(scr_p2, file="scrubOut_p2.RData")
#save(scr_p3, file="scrubOut_p3.RData")
#save(scr_p4, file="scrubOut_p4.RData")

load("~/scrubOut_p1.RData")
load("~/scrubOut_p2.RData")
load("~/scrubOut_p3.RData")
load("~/scrubOut_p4.RData")
```

## Examine output
### 2022b Plate 1
```{r}
# Estimated contamination level
scr_p1$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1$p*100))  #0.74% contamination
```
### 2022b Plate 2
```{r}
scr_p2$p %>% boxplot()
```
```{r}
median(100-(scr_p2$p*100))  #1.98% contamination
```
### 2022b Plate 3
```{r}
scr_p3$p %>% boxplot()
```
```{r}
median(100-(scr_p3$p*100))  #2.68% contamination
```
### 2022b Plate 4
```{r}
scr_p4$p %>% boxplot()
```
```{r}
median(100-(scr_p4$p*100))  #5.16% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1
```{r}
#  index = 1 to 38904 ASVs in this dataset
scr_p1$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 2
```{r}
scr_p2$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 3
```{r}
scr_p3$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 4
```{r}
scr_p4$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot() 
```

## Estimate well-to-well leakage
### 2022b Plate 1
```{r}
scr_p1$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.10% well-to-well leakage in plate 1
```
### 2022b Plate 2
```{r}
scr_p2$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p2$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.34% well leakage in plate 2
```
### 2022b Plate 3
```{r}
scr_p3$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p3$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.25% well leakage in plate 3
```
### 2022b Plate 4
```{r}
scr_p4$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p4$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.90% well leakage in plate 4
```

```{r}
decon_p1 <- scr_p1$decontaminated_samples
decon_p2 <- scr_p2$decontaminated_samples
decon_p3 <- scr_p3$decontaminated_samples
decon_p4 <- scr_p4$decontaminated_samples
write.csv(decon_p1, "2022b_ScrubDecontaminatedSamples_p1_pcr.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p2, "2022b_ScrubDecontaminatedSamples_p2_pcr.csv")
write.csv(decon_p3, "2022b_ScrubDecontaminatedSamples_p3_pcr.csv")
write.csv(decon_p4, "2022b_ScrubDecontaminatedSamples_p4_pcr.csv")

#decon_p1 <- read.csv("~/2022b_ScrubDecontaminatedSamples_p1_pcr.csv", row.names = 1) %>% as.matrix()
#decon_p2 <- read.csv("~/2022b_ScrubDecontaminatedSamples_p2_pcr.csv", row.names = 1) %>% as.matrix()
#decon_p3 <- read.csv("~/2022b_ScrubDecontaminatedSamples_p3_pcr.csv", row.names = 1) %>% as.matrix()
#decon_p4 <- read.csv("~/2022b_ScrubDecontaminatedSamples_p4_pcr.csv", row.names = 1) %>% as.matrix()
```

## Phase2 SCRuB - Extraction Neg Ctrl
### Run SCRuB
* Running SCRuB with spatial (well location) information. Done by plate.
```{r}
#scrub DNA plates (individually) by DNA extraction controls
scr_p1_ex <- SCRuB(mat_p1, meta_p1, c("Ex-Neg-Ctrl") )
scr_p2_ex <- SCRuB(mat_p2, meta_p2, c("Ex-Neg-Ctrl") ) 
scr_p3_ex <- SCRuB(mat_p3, meta_p3, c("Ex-Neg-Ctrl") )
scr_p4_ex <- SCRuB(mat_p4, meta_p4, c("Ex-Neg-Ctrl") )
```
## Saving and loading Scrubbed files
```{r}
#save(scr_p1_ex, file="scrubOut_p1_ex.RData")
#save(scr_p2_ex, file="scrubOut_p2_ex.RData")
#save(scr_p3_ex, file="scrubOut_p3_ex.RData")
#save(scr_p4_ex, file="scrubOut_p4_ex.RData")

load("~/scrubOut_p1_ex.RData")
load("~/scrubOut_p2_ex.RData")
load("~/scrubOut_p3_ex.RData")
load("~/scrubOut_p4_ex.RData")
```

## Examine output
### 2022b Plate 1
```{r}
# Estimated contamination level
scr_p1_ex$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1_ex$p*100))  #4.41% 
```
### 2022b Plate 2
```{r}
scr_p2_ex$p %>% boxplot()
```
```{r}
median(100-(scr_p2_ex$p*100))  #1.75% contamination
```
### 2022b Plate 3
```{r}
scr_p3_ex$p %>% boxplot()
```
```{r}
median(100-(scr_p3_ex$p*100))  #5.36% contamination
```
### 2022b Plate 4
```{r}
scr_p4_ex$p %>% boxplot()
```
```{r}
median(100-(scr_p4_ex$p*100))  #3.27% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1
```{r}
#  index = 1 to 7781 ASVs in this dataset
scr_p1_ex$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 2
```{r}
scr_p2_ex$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 3
```{r}
scr_p3_ex$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```
### 2022b Plate 4
```{r}
scr_p4_ex$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
```

## Estimate well-to-well leakage
### 2022b Plate 1
```{r}
scr_p1_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.68% well-to-well leakage in plate 1
```
### 2022b Plate 2
```{r}
scr_p2_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p2_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.25% well leakage in plate 2
```
### 2022b Plate 3
```{r}
scr_p3_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p3_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.75% well leakage in plate 3
```
### 2022b Plate 4
```{r}
scr_p4_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p4_ex$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.55% well leakage in plate 4
```

```{r}
decon_p1_ex <- scr_p1_ex$decontaminated_samples
decon_p2_ex <- scr_p2_ex$decontaminated_samples
decon_p3_ex <- scr_p3_ex$decontaminated_samples
decon_p4_ex <- scr_p4_ex$decontaminated_samples
write.csv(decon_p1_ex, "2022b_ExNegCtrl_ScrubDecontaminatedSamples_p1.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p2_ex, "2022b_ExNegCtrl_ScrubDecontaminatedSamples_p2.csv")
write.csv(decon_p3_ex, "2022b_ExNegCtrl_ScrubDecontaminatedSamples_p3.csv")
write.csv(decon_p4_ex, "2022b_ExNegCtrl_ScrubDecontaminatedSamples_p4.csv")

#decon_p1_ex <- read.csv("~/2022b_ExNegCtrl_ScrubDecontaminatedSamples_p1.csv", row.names = 1) %>% as.matrix()
#decon_p2_ex <- read.csv("~/2022b_ExNegCtrl_ScrubDecontaminatedSamples_p2.csv", row.names = 1) %>% as.matrix()
#decon_p3_ex <- read.csv("~/2022b_ExNegCtrl_ScrubDecontaminatedSamples_p3.csv", row.names = 1) %>% as.matrix()
#decon_p4_ex <- read.csv("~/2022b_ExNegCtrl_ScrubDecontaminatedSamples_p4.csv", row.names = 1) %>% as.matrix()
```

# Output post-SCRuB 
```{r}
# create main datsets - putting plates back together again - combining PCR and Ex SCRuB
mat_final <- rbind(decon_p1, decon_p2, decon_p3, decon_p4)
ps_curated_0 <- subset_samples(ps_22b, sample_names(ps_22b) %in% rownames(mat_final))
ps_curated <- ps_curated_0
mat_table_final <- otu_table(mat_final, taxa_are_rows = FALSE)
otu_table(ps_curated) <- mat_table_final

mat_final_ex <- rbind(decon_p1_ex, decon_p2_ex, decon_p3_ex, decon_p4_ex)
ps_curated_0_ex <- subset_samples(ps_22b, sample_names(ps_22b) %in% rownames(mat_final_ex))
ps_curated_ex <- ps_curated_0_ex
mat_table_final_ex <- otu_table(mat_final_ex, taxa_are_rows = FALSE)
otu_table(ps_curated_ex) <- mat_table_final_ex
```
## Seeing how ASV reads change per sample before and after scrubbing
### Plate 1
```{r}
ps_curated_0@otu_table["BM-290",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-290",1:10]   #post SCRuB
```

### Plate 2
```{r}
ps_curated_0@otu_table["BM-35",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-35",1:10]   #post SCRuB
```

### Plate 3
```{r}
ps_curated_0@otu_table["BM-375",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-375",1:10]   #post SCRuB
```

### Plate 4
```{r}
ps_curated_0@otu_table["BM-250b",1:10] #prior SCRuB
```
```{r}
ps_curated@otu_table["BM-250b",1:10]   #post SCRuB
```

### Plate 1 Ex
```{r}
ps_curated_0_ex@otu_table["BM-290",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-290",1:10]   #post SCRuB
```

### Plate 2 Ex
```{r}
ps_curated_0_ex@otu_table["BM-35",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-35",1:10]   #post SCRuB
```

### Plate 3 Ex
```{r}
ps_curated_0_ex@otu_table["BM-375",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-375",1:10]   #post SCRuB
```

### Plate 4 Ex
```{r}
ps_curated_0_ex@otu_table["BM-250b",1:10] #prior SCRuB
```
```{r}
ps_curated_ex@otu_table["BM-250b",1:10]   #post SCRuB
```

```{r}
row_sums <- rowSums(ps_curated_0@otu_table) #prior scrub
print(row_sums)
```
```{r}
row_sums1 <- rowSums(ps_curated@otu_table) #post scrub
print(row_sums1)
```

```{r}
row_sums2 <- rowSums(ps_curated_0_ex@otu_table) #prior scrub
print(row_sums2)
```
```{r}
row_sums3 <- rowSums(ps_curated_ex@otu_table) #post scrub
print(row_sums3)
```

```{r}
# filter out taxa reduced to 0 reads (19495 ASVs, leaves 19409 out of 38904)
ps_curated <- filter_taxa(ps_curated, function(x) sum(x) > 0, TRUE)
#ps_22b
ps_curated

# filter out taxa reduced to 0 reads (19531 ASVs, leaves 19373 out of 38904)
ps_curated_ex <- filter_taxa(ps_curated_ex, function(x) sum(x) > 0, TRUE)
ps_curated_ex
```

### Remove positive control samples
```{r}
ps_curated <- prune_samples(sample_names(ps_curated) != "P1-Pos-Ctrl", ps_curated)
ps_curated <- prune_samples(sample_names(ps_curated) != "P2-Pos-Ctrl", ps_curated)
ps_curated <- prune_samples(sample_names(ps_curated) != "P3-Pos-Ctrl", ps_curated)
ps_curated <- prune_samples(sample_names(ps_curated) != "P4-Pos-Ctrl", ps_curated)
ps_curated@sam_data
```
```{r}
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P1-Pos-Ctrl", ps_curated_ex)
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P2-Pos-Ctrl", ps_curated_ex)
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P3-Pos-Ctrl", ps_curated_ex)
ps_curated_ex <- prune_samples(sample_names(ps_curated_ex) != "P4-Pos-Ctrl", ps_curated_ex)
ps_curated_ex@sam_data
```

## Merging SCRuBBED 2022 plates
### Merging Ex Neg Ctrl and PCR Neg Ctrl SCRuB
```{r}
#this will merge the samples with the same name - they will be added together
ps_22b_merge <- merge_phyloseq(ps_curated, ps_curated_ex)
ps_22b_merge
sample_data(ps_22b_merge)
```

## 2022b Combined SCRuB
```{r}
## Running SCRuB of both Ex Neg and PCR Neg Ctrl at one time
## According to SCRuB tutorial, run SCRuB of first instance of contamination first

#scrub DNA plates (individually) by DNA extraction and PCR controls
scr_p1_comb <- SCRuB(mat_p1, meta_p1, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") )
scr_p2_comb <- SCRuB(mat_p2, meta_p2, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") ) 
scr_p3_comb <- SCRuB(mat_p3, meta_p3, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") )
scr_p4_comb <- SCRuB(mat_p4, meta_p4, control_order =  c("Ex-Neg-Ctrl", "PCR-Neg-Ctrl") )
```

## Saving and loading Scrubbed files
```{r}
#save(scr_p1_comb, file="scrubOut_p1_comb.RData")
#save(scr_p2_comb, file="scrubOut_p2_comb.RData")
#save(scr_p3_comb, file="scrubOut_p3_comb.RData")
#save(scr_p4_comb, file="scrubOut_p4_comb.RData")

load("~/scrubOut_p1_comb.RData")
load("~/scrubOut_p2_comb.RData")
load("~/scrubOut_p3_comb.RData")
load("~/scrubOut_p4_comb.RData")
```

## Examine output
### 2022b Plate 1
```{r}
# Estimated contamination level
scr_p1_comb$p %>% boxplot()  #read this as closer to 1 = less contaminated
```
```{r}
median(100-(scr_p1_comb$p*100))  #4.81% 
```
### 2022b Plate 2
```{r}
scr_p2_comb$p %>% boxplot()
```
```{r}
median(100-(scr_p2_comb$p*100))  #3.74% contamination
```
### 2022b Plate 3
```{r}
scr_p3_comb$p %>% boxplot()
```
```{r}
median(100-(scr_p3_comb$p*100))  #6.02% contamination
```
### 2022b Plate 4
```{r}
scr_p4_comb$p %>% boxplot()
```
```{r}
median(100-(scr_p4_comb$p*100))  #2.73% contamination
```

## Estimated realtive abundance of the contamination community
### 2022b Plate 1
```{r}
#  index = 1 to 7781 ASVs in this dataset
scr_p1_comb$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot()
scr_p1_comb$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```
### 2022b Plate 2
```{r}
scr_p2_comb$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot()
scr_p2_comb$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```
### 2022b Plate 3
```{r}
scr_p3_comb$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot()
scr_p3_comb$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```
### 2022b Plate 4
```{r}
scr_p4_comb$inner_iterations$`Ex-Neg-Ctrl`$gamma %>% plot() 
scr_p4_comb$inner_iterations$`PCR-Neg-Ctrl`$gamma %>% plot()
```

## Estimate well-to-well leakage
### 2022b Plate 1
```{r}
scr_p1_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p1_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.68% well-to-well leakage in plate 1
scr_p1_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p1_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.10% well-to-well leakage in plate 1
```
### 2022b Plate 2
```{r}
scr_p2_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p2_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.24% well leakage in plate 2
scr_p2_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p2_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.34% well leakage in plate 2
```
### 2022b Plate 3
```{r}
scr_p3_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p3_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.75% well leakage in plate 3
scr_p3_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p3_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.20% well leakage in plate 3
```
### 2022b Plate 4
```{r}
scr_p4_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha[, ncol(scr_p4_comb$inner_iterations$`Ex-Neg-Ctrl`$alpha)]
# 0.10% well leakage in plate 4
scr_p4_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha[, ncol(scr_p4_comb$inner_iterations$`PCR-Neg-Ctrl`$alpha)]
# 0.74% well leakage in plate 4
```

```{r}
decon_p1_comb <- scr_p1_comb$decontaminated_samples
decon_p2_comb <- scr_p2_comb$decontaminated_samples
decon_p3_comb <- scr_p3_comb$decontaminated_samples
decon_p4_comb <- scr_p4_comb$decontaminated_samples
write.csv(decon_p1_comb, "2022b_combNegCtrl_ScrubDecontaminatedSamples_p1.csv") #save each of these as CSV file to then be able to put back together
write.csv(decon_p2_comb, "2022b_combNegCtrl_ScrubDecontaminatedSamples_p2.csv")
write.csv(decon_p3_comb, "2022b_combNegCtrl_ScrubDecontaminatedSamples_p3.csv")
write.csv(decon_p4_comb, "2022b_combNegCtrl_ScrubDecontaminatedSamples_p4.csv")

#decon_p1 <- read.csv("~/2022b_combNegCtrl_ScrubDecontaminatedSamples_p1.csv", row.names = 1) %>% as.matrix()
#decon_p2 <- read.csv("~/2022b_combNegCtrl_ScrubDecontaminatedSamples_p2.csv", row.names = 1) %>% as.matrix()
#decon_p3 <- read.csv("~/2022b_combNegCtrl_ScrubDecontaminatedSamples_p3.csv", row.names = 1) %>% as.matrix()
#decon_p4 <- read.csv("~/2022b_combNegCtrl_ScrubDecontaminatedSamples_p4.csv", row.names = 1) %>% as.matrix()
```

# Output post-SCRuB 
```{r}
# create main datsets - putting plates back together again - combining PCR and Ex SCRuB
mat_final_comb <- rbind(decon_p1_comb, decon_p2_comb, decon_p3_comb, decon_p4_comb)
ps_curated_0_comb <- subset_samples(ps_22b, sample_names(ps_22b) %in% rownames(mat_final_comb))
ps_curated_comb <- ps_curated_0_comb
mat_table_final_comb <- otu_table(mat_final_comb, taxa_are_rows = FALSE)
otu_table(ps_curated_comb) <- mat_table_final_comb
```

### Plate 1 Combined
```{r}
ps_curated_0_comb@otu_table["BM-290",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-290",1:10]   #post SCRuB
```

### Plate 2 Combined
```{r}
ps_curated_0_comb@otu_table["BM-35",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-35",1:10]   #post SCRuB
```

### Plate 3 Combined
```{r}
ps_curated_0_comb@otu_table["BM-375",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-375",1:10]   #post SCRuB
```

### Plate 4 Combined
```{r}
ps_curated_0_comb@otu_table["BM-250b",1:10] #prior SCRuB
```
```{r}
ps_curated_comb@otu_table["BM-250b",1:10]   #post SCRuB
```

```{r}
row_sums12 <- rowSums(ps_curated_0_comb@otu_table) #prior scrub
print(row_sums12)
```
```{r}
row_sums13 <- rowSums(ps_curated_comb@otu_table) #post scrub
print(row_sums13)
```

```{r}
# filter out taxa reduced to 0 reads (25637 ASVs, leaves 13267 out of 38904)
ps_curated_comb <- filter_taxa(ps_curated_comb, function(x) sum(x) > 0, TRUE)
#ps_22b
ps_curated_comb
```

### Removing Positive Controls
```{r}
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P1-Pos-Ctrl", ps_curated_comb)
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P2-Pos-Ctrl", ps_curated_comb)
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P3-Pos-Ctrl", ps_curated_comb)
ps_curated_comb <- prune_samples(sample_names(ps_curated_comb) != "P4-Pos-Ctrl", ps_curated_comb)
ps_curated_comb@sam_data
```

## Saving Files
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(ps_curated_comb))
names(dna) <- taxa_names(ps_curated_comb)
ps_22b_scrub <- merge_phyloseq(ps_curated_comb, dna)
taxa_names(ps_22b_scrub) <- paste0("ASV", seq(ntaxa(ps_22b_scrub)))
ps_22b_scrub

saveRDS(ps_22b_scrub,file.path(path,"2022b_nextseq.ps.rds"))
```

```{r}
# saving files
scrub_meta <- (as(sample_data(ps_22b_scrub), "data.frame"))
scrub_taxa <- (as(otu_table(ps_22b_scrub), "matrix"))
# # write out data
write.csv(scrub_meta, "2022b_metadata_postSCRuB_nextseq_comb.csv")
write.csv(scrub_taxa, "2022b_taxatable_postSCRuB_nextseq_comb.csv")
```

```{r}
sessionInfo()
```





###### end